---
title: "Getting Started with R"
subtitle: "Basic Operations and a First ggplot2 Chart"
author: "Alexa Bueno"
date: today
format:
  revealjs:
    theme: simple            # built-in theme, try moon, sky, or serif too
    slide-number: true       # shows slide numbers in the corner
    transition: slide        # animation between slides
    code-line-numbers: true
    embed-resources: true    # packs everything into one .html file
---

# The Basics

## Why R?

::: {.incremental}
- Free and open source
- Built for statistics and data visualization
- Thousands of packages for almost any analysis
- Works with Quarto, so code, charts, and writing live in one file
:::

::: notes
Speaker notes go here. Press S during the presentation to see them.
:::

## R as a Calculator

```{r}
#| echo: true
1 + 1
2 * 2
3^3   # ^ is the exponent operator in R
```

## Storing Values

:::: {.columns}

::: {.column width="50%"}
```{r}
#| echo: true
message <- "Hello world!"
print(message)

number <- 7
number
```
:::

::: {.column width="50%"}
```{r}
#| echo: true
max(2, 5, 90, 30)
min(2, 5, 90, 30)
```
:::

::::

`<-` is the assignment operator. R users prefer it over `=`.

# Visualizing Data

## Building a Density Plot

```{r}
#| echo: true
#| eval: false
#| code-line-numbers: "|1|3|5-6|7-11"
library(ggplot2)

theme_set(theme_classic())

g <- ggplot(mpg, aes(cty))
g + geom_density(aes(fill = factor(cyl)), alpha = 0.8) +
  labs(title = "Density plot",
       subtitle = "City Mileage Grouped by Number of Cylinders",
       caption = "Source: mpg dataset",
       x = "City Mileage",
       fill = "# Cylinders")
```

::: notes
Clicking through highlights each part of the code step by step.
:::

## The Result

```{r}
#| echo: false
#| fig-width: 9
#| fig-height: 5
library(ggplot2)
theme_set(theme_classic())

g <- ggplot(mpg, aes(cty))
g + geom_density(aes(fill = factor(cyl)), alpha = 0.8) +
  labs(title = "Density plot",
       subtitle = "City Mileage Grouped by Number of Cylinders",
       caption = "Source: mpg dataset",
       x = "City Mileage",
       fill = "# Cylinders")
```

## Takeaways

::: {.incremental}
- Cars with fewer cylinders get better city mileage
- ggplot2 builds charts in layers connected with `+`
- Because the chart is generated from code, anyone can reproduce it
:::